---
title: "ggplot intro"
author: "jmittermeier"
date: "July 13, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Plan

We'll load libraries and learn ggplot2.

Here we write in Markdown.

# hashtags mean headers.

```{r name your rchunks in the curly brackets}
# here we write R.
# hashtags mean comment.
```

``` {r}
# loading libraries

library(dplyr)
library(gapminder)
library(ggplot2) # install.packages('ggplot2')

## subset the data
g <- gapminder %>%
  filter(year == 2007) %>%
  mutate(pop_m = pop/1e6)
```

# My first plot

a simple geom_point

``` {r}

## ggplot(data, aes(what's x, what's y)) + what_kind_of_plot()
s <- ggplot(g, aes(x = gdpPercap, y = lifeExp)) +
  geom_point()
s

```
  
# My second plot

a simple geom_point

``` {r}

## ggplot(data, aes(what's x, what's y)) + what_kind_of_plot()
s2 = s + aes(size = pop_m)
s2

s3 = s2 + aes(color = continent)
s3

``` 

# A new chunk with everything added together
This makes things more streamlined
``` {r}

## ggplot(data, aes(what's x, what's y)) + what_kind_of_plot()
s <- ggplot(g, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() + aes(size = pop_m) + aes(color = continent) +
  ggtitle('Health and Wealth of Nations for 2007') +
  xlab('GDP per capita (USD/year)') +
  ylab('Life Expectancy (years)') +
  scale_colour_discrete('Continent') +
  scale_size_continuous('pop_m')
s

``` 

## Make a box plot

This is my boxplot of continent and life expectancy.  This is done using 'g', which is a subset of 2007
``` {r}

## ggplot(data, aes(what's x, what's y)) + what_kind_of_plot()
b <- ggplot(g, aes(x = continent, y = lifeExp)) + geom_boxplot() +
  aes(fill = continent) +
  theme(legend.position = 'none')
b

```  

## GoogleVis interactive figure

``` {r, results='asis'}

# don't forget the 'asis' bit at the start of your r code in this case

suppressPackageStartupMessages(library(googleVis)) #install.packages('googleVis')
#use the suppress packages to remove the initial start up message

#code copied from course pages
op <- options(gvis.plot.tag='chart')

m = gvisMotionChart(
  gapminder %>%
    mutate(
      pop_m = pop / 1e6,
      log_gdpPercap = log(gdpPercap)), 
  idvar='country', 
  timevar='year', 
  xvar='log_gdpPercap', 
  yvar='lifeExp', 
  colorvar='continent',
  sizevar='pop_m')
plot(m)

```  
  
  
  
  
  
  
  